---
title: "Data Wrangling"
output: pdf_document
---

# Data Wrangling
This document showcases how the datasets are wrangled to improve the depth of the variables.

### Load tidyverse
The tidyverse package was loaded. This package will make it easy to wrangle and clean data.
```{r}
## This is the function for loading the tidyverse package
library(tidyverse)
```

### Load datasets
The needed datasets are loaded in the column below.
```{r, load-data}
econ_free <- read.csv("~/Global-Analysis/Economic_Freedom_data.csv")
df_gap <- read.csv("~/Global-Analysis/new_gapminder.csv")
```

## Combining Datasets

### Transforming the gapminder dataset
The cell below shows how gapminder was transformed to create a new column.This column will be used to uniquely identify each country during each year.
```{r}
## The data type for the year column was transformed into character
df_gap$char <- as.character(df_gap$year)
## Another new column was created to duplicate the country information
df_gap$newcountry <- df_gap$country

## Then, the new country column was combined with the year to
## create a new column for joining 
df_gap <- df_gap %>% 
  unite("country_year", c(newcountry, char), sep = ", ")
```

### Checking out gapminder dataset

```{r}
head(df_gap)
```
### Transforming the economic indicators index dataset

The cell below shows how economic indicators' dataset were transformed to create a new column.This column will be used to uniquely identify each country during each year.

```{r}
## The same process that was used to create a country year column
## was repeated below
econ_free$char <- as.character(econ_free$Year)
econ_free$newcountry <- econ_free$Country
econ_free <- econ_free %>% 
  unite("country_year", c(newcountry, char), sep = ", ")
```

### Checking out the dataset

```{r}
head(econ_free)
```
### Joining the two datasets
The cell below was used to join the two columns using the newly formed identifier country_year.
```{r}
## The two datasets were joined together using the column 
## country_year to form a new dataset
df_world <- inner_join(df_gap, econ_free, by = "country_year")

## Some columns were left out below
df_world <- df_world %>% 
  select(-c(Country, Year))
```

### Saving the datasets

```{r}
## The dataset was saved into a new .csv file
write.csv(df_world, "~/Global-Analysis/New_world.csv")
```


