---
title: "Exploratory Data Analysis"
output: pdf_document
---

# Exploratory Data Analysis

This is a document for exploring the global indicator dataset that was extracted from the combination of numerous datasets

### Loading packages and dataset

The tidyverse package was loaded below.

```{r}
library(tidyverse)
```

The dataset is loaded below

```{r}
df_world <- read.csv("~/Global-Analysis/World_data.csv")
```

### Data Comprehension

The datasets was reviewed below to improve our understanding of the dataset

```{r}
head(df_world)
colnames(df_world)
```

Structure of the dataset

```{r}
str(df_world)
```

Renaming certain columns

```{r}
df_world <- df_world %>%
  rename(c(Economic_Freedom_Index = Economic.Freedom.Index, 
           Life_Expectancy = Life.Expectancy,
           Government_Integrity = Government.Integrity,
           Government_Spending = Government.Spending,
           Property_Rights = Property.Rights,
           Fiscal_Health = Fiscal.Health,
           Business_Freedom = Business.Freedom,
           Monetary_Freedom = Monetary.Freedom,
           Trade_Freedom = Trade.Freedom,
           Investment_Freedom = Investment.Freedom,
           Financial_Freedom = Financial.Freedom,
           Income_Index = Income.Index,
           Tax_Burden = Tax.Burden,
           Labor_Freedom = Labor.Freedom,
           Expected_schooling = Expected.years.of.Schooling,
           Education_Index = Education.Index,
           Human_Development_Index = Human.Development.Index,
           Judicial_Effectiveness = Judicial.Effectiveness,
           Free_Market_Class = Free.Market.Class
           ))
```

Filtering out information

```{r}
df_nigeria <- df_world %>% 
  filter(Country == 'Nigeria') %>%
  select(Business_Freedom, Year)
df_nigeria
```

## Charts

Scatter plot function

```{r}
## Human Development Index vs Economic Freedom Index
scatter1 <- function(df, year) {
  df %>% 
    filter(Year == year) %>% 
    ggplot() + 
    geom_point(aes(x = Economic_Freedom_Index, 
                     y = Human_Development_Index, 
                   color = Continent), na.rm = TRUE) +
    labs(title = sprintf("Human Development Index vs Economic Freedom Index in %d", year), 
         x = "Economic Freedom Index",
         y = "Human Development Index")
    
}
## Education Index vs Economic Freedom Index
scatter2 <- function(df, year) {
  df %>% 
    filter(Year == year) %>% 
    ggplot() + 
    geom_point(aes(x = Economic_Freedom_Index, 
                   y = Education_Index, 
                   color = Continent), na.rm = TRUE) +
    labs(title = sprintf("Education Index vs Economic Freedom Index in %d", year), 
         x = "Economic Freedom Index",
         y = "Education Index")
}
## Life Expectancy vs Economic Freedom Index
scatter3 <- function(df, year) {
  df %>% 
    filter(Year == year) %>% 
    ggplot() + 
    geom_point(aes(x = Economic_Freedom_Index, 
                     y = Life_Expectancy, 
                   color = Continent), na.rm = TRUE) +
    labs(title = sprintf("Life Expectancy vs Economic Freedom Index in %d", year), 
         x = "Economic Freedom Index",
         y = "Life Expectancy")
}
## Government Integrity vs Government Spending
scatter4 <- function(df, year) {
  df %>% 
    filter(Year == year) %>% 
    ggplot() + 
    geom_point(aes(x = Government_Spending, 
                     y = Government_Integrity, 
                   color = Continent), na.rm = TRUE) +
    labs(title = sprintf("Government Integrity vs Government Spending in %d", year), 
         x = "Government Spending",
         y = "Government Integrity")
}
```

Human Development Index vs Economic Freedom Index

```{r, fig.width = 9, fig.height = 9}
scatter1(df_world, 2019)
```

Education Index vs Economic Freedom Index

```{r, fig.width = 9, fig.height = 9}
scatter2(df_world, 2019)
```

Life Expectancy vs Economic Freedom Index

```{r, fig.width = 9, fig.height = 9}
scatter3(df_world, 2019)
```
Government Integrity vs Government Spending
```{r, fig.width = 12, fig.height=9}
scatter4(df_world, 2019)
```

Kernel Density Estimation Plot

Human Development Index

```{r, fig.width = 12, fig.height=9}
ggplot(df_world) +
  geom_density(aes(x = Human_Development_Index, 
                    color = as.character(Year)),
                show.legend = NA) +
  labs(title = "How Human Development Index has changed over time")
```

Economic Freedom Index

```{r, fig.width = 12, fig.height=9}
ggplot(df_world) +
  geom_density(aes(x = Economic_Freedom_Index, 
                    color = as.character(Year)),
                show.legend = NA) +
  labs(title = "How Economic Freedom Index has changed over time")
```

Box Plot

Continent Disparities for Human Development Index in 2019

```{r, fig.width = 12, fig.height = 9}
df_world %>% 
  filter(Year == 2019) %>% 
  ggplot() + 
  geom_boxplot(aes(x = Continent, y = Human_Development_Index,
                   color = Continent)) +
  labs(title = "How continents differ based on Human Development Index in 2019")
```

Continent Disparities for Economic Freedom Index in 2019

```{r, fig.width = 12, fig.height = 9}
df_world %>% 
  filter(Year == 2019) %>% 
  ggplot() + 
  geom_boxplot(aes(x = Continent, y = Economic_Freedom_Index,
                   color = Continent)) +
  labs(title = "How continents differ based on Economic Freedom Index in 2019", 
       y = "Economic Freedom Index")
```

Bar Chart

Average life expectancy across continents

```{r, fig.width= 12, fig.height=9}
df_world %>% 
  filter(Year == 2019) %>% 
  group_by(Continent) %>% 
  summarize(mean_life = mean(Life_Expectancy, na.rm = TRUE)) %>% 
  ggplot() +
  geom_bar(aes(x = Continent, y = mean_life, 
               fill = Continent), stat = "identity") +
  labs(title = "Average life expectancy across continents", 
       y = "Average life expectancy") +
  theme_classic()
```

Enhanced scatter plot for Human Development Index vs Economic Freedom Index

```{r, fig.width = 12, fig.height = 8}
  df_world %>% 
    filter(Year == 2019) %>% 
    ggplot() + 
    geom_point(aes(x = Economic_Freedom_Index, 
                     y = Human_Development_Index, 
                   color = Continent), na.rm = TRUE) +
    facet_grid(Continent ~ Free_Market_Class) +
    labs(title = sprintf("Human Development Index vs Economic Freedom Index in %d", 2019), 
         x = "Economic Freedom Index",
         y = "Human Development Index") 
```

